---
title: "Precision-Recall Curves"
author: "Gaurav Kandoi"
date: "February 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Precision-Recall Curve

This uses the PRROC library

```{r}
#install.packages("PRROC")
library(PRROC)
```

Set a random seed so that everyone gets the same results with the example script.  
First we generate 1000 random "Scores" between 0 and 1. These will be the scores from your prediction for every edge.  
The "True_Labels" are the known labels (0 or 1) for every gene pair. 0 means no edge (negative) and 1 means the presence of an edge (positive).  

```{r}
set.seed(1)
Scores <- runif(1000)
True_Labels <- sample(c(0,1), size = 1000, replace = T)
```

Next we use the "pr.curve" function from PRROC to calcuate the precision-recall values. The function also calculate the area under the precision recall curve.    
The function requires your true labels and prediction scores.  
Please note that the value i in the true labels and prediction scores should correspond to the same gene pair.  
Next we plot the precision recall curve using the base plot function.  

```{r}
pr <- pr.curve(scores.class0 = Scores, weights.class0 = True_Labels, curve = T)
pr
plot(pr)
```

System information
```{r}
R.Version()
search()
```
